Management induced changes of soil organic carbon on global croplands

### 2022-02-28 SOCbudgetPaper/SOCbudget.tex
remove copyright statement fix typo in reference
Major update of finalized Paper - Draft 2

%% Journal abbreviations (please use the same for discussion papers and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% ASTRA Proceedings (ap)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

@@

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
%
\usepackage{algorithmic}
\usepackage{algorithm}

\begin{document}

@@

\begin{abstract}
Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth. The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere. SOC has been depleted by anthropogenic land cover change and agricultural management. However, the latter has so far not been well represented in global carbon stock assessments. While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale. Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils. We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975. Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity. We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
\end{abstract}

\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}

@@

Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth's soil, is the largest terrestrial organic carbon pool. It exceeds the carbon in the atmospheric and vegetation pools multiple times \citep{batjes_total_1996}. Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere \citep{friedlingstein_global_2019, minasny_soil_2017}. The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile \citep{batjes_total_1996, sanderman_soil_2017}.

Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time. Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content \citep{batjes_harmonized_2016, hengl_soilgrids250m_2017, fao_global_2018}. However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis. Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.

The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state. Sanderman et al.~\citeyearpar{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter \citep{lal_world_2001}.

Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions \citep{friedlingstein_global_2019}. While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management. Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest \citep{strassmann_simulating_2008, betts_climate_2015}. BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management \citep{friedlingstein_global_2019, houghton_carbon_2012, hansis_relevance_2015}.

Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance \citep[e.g.~][]{bondeau_modelling_2007, lindeskog_implications_2013}. Pugh et al.~\citeyearpar{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data. They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons \citep{portmann_mirca2000global_2010} multiple cropping systems \citep{waha_multiple_2020}, irrigation systems \citep{jagermeyr_water_2015}) or as stylized scenarios \citep[e.g.~][\citet{lutz_simulating_2019}]{pugh_simulated_2015}.

More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.~on tillage systems, see \citep{porwollik_generating_2018}, \citep{prestele_spatially_2018}) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019}, \citep{minoli_modelling_2019}), but have --- to our knowledge --- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.

Field-scale models \citep[\citet{coleman_simulating_1997}, \citet{smith_estimating_2010}, \citet{taghizadeh-toosi_c-tool_2014}]{del_grosso_simulated_2001} are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.

The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management. To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We simulate SOC stocks, dynamics and the SOC debt for 1975--2010. Using a scenario analysis, we decompose the contribution of different management activities.
\newpage

@@

In Sect. \ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 \citep{calvo_buendia_ipcc_2019}. We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}. In Sect. \ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019}.
Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices \citep[for model code see][]{bodirsky_mrcommons_2020}. In Sect. \ref{sec:scenarios} we define the scenarios used to complement our historic model results.

\hypertarget{sec:carbonbudget}{%
\subsection{SOC stocks and stock changes following the steady-state method}\label{sec:carbonbudget}}

Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.~4 \citep{calvo_buendia_ipcc_2019}; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon (\(SOC\)) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010. We restrict our analysis to the first 0-30 cm of the soil profile. Moreover, we assume the current \(SOC\) state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
Therefore, we take the following three steps for each year of our simulation period:
(1) We calculate annual land-use type-specific steady states and decay rates for \(SOC\) stocks (Sect. \ref{sec:steadystates});
(2) we account for land conversion by transferring \(SOC\) from and to natural vegetation (Sect. \ref{sec:carbontransfer}),
(3) we estimate \(SOC\) stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \ref{sec:totalsoc}).
To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \ref{sec:initsoc}).

@@

In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{\mathrm{eq}}\) are given by
\begin{equation}
SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
\label{eq:inoutflow}
\end{equation}
with \(C^{\textrm{in}}\) being the carbon inputs to the soil, \(k\) denotes the soil organic carbon decay rate. This equation is valid for all grid cells \(i\) and all years \(t\). We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model \citep{parton_analysis_1987}. Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).

Carbon inputs for cropland are below- and above-ground residues left or returned to the field (see Sect. \ref{sec:residues}) and manure inputs (see Sect. \ref{sec:livstmanure}); for natural vegetation litterfall including fine root turnover \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil. Following the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content. For each component the sum of all carbon input sources is allocated to the respective \(SOC\) sub-pools via transfer coefficients. This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools. Data sources for all considered carbon inputs as well as for lignin and nitrogen content are listed in Table \ref{tab:datasourceinputs}.

  \begin{table*}[h]
  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
  \tophline
  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
  \middlehline
  \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                      \citet{faostat_faostat_2016}, \\
                                                                      \citet{schaphoff_lpjml4_2018},\\
                                                                      \citet{weindl_livestock_2017} \end{minipage}} & 
                                                      \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                      LG:C generic values according to Table 5.5B,\\
                                                                      5.5C from IPCC \citep{calvo_buendia_ipcc_2019},\\
                                                                      crop-specific N:C from \citet{bodirsky_n2o_2012}
                                                                      \end{minipage}} \\
                            & below-ground residues, &  &  \\
                            & manure                &  &  \\ 
                            \hline
  Natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
                            & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
                                IPCC \citep{calvo_buendia_ipcc_2019} and\\ 
                                CENTURY (\citep{century_model_2000}) \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \label{tab:datasourceinputs}
 \belowtable{}
 \end{table*}

The sub-pool specific decay rates \(k_{sub}\) are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space (\(i\)) and time (\(t\)). Following the steady-state method \citep{calvo_buendia_ipcc_2019}, we consider temperature (\(temp\)), water (\(wat\)), sand fraction (\(sf\)) and tillage (\(till\)) effects to account for spatial variation of decay rates. Thus, \(k_{sub}\) are given by

@@

For natural vegetation, we assume rainfed and non-tilled conditions, whereas for cropland, we distinguish the effect of different tillage (see Sect. \ref{sec:tillage}) and irrigation (see Sect. \ref{sec:irrigation}) practices on decay rates. We calculated area weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation pratices. Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equation 5.0B--5.0F in Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.

 \begin{table*}[h]
 \caption{Type and data sources for carbon inputs to different land-use types}
 \begin{tabular}{l l l l}
 \tophline
  land-use types   & type of decay driver & parameter use to represent driver & data source \\
 \middlehline
 \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
                                     & \citet{hengl_soilgrids250m_2017} \\
                      \cline{2-4}

                      & Mircobial activity & air temperature & \citet{harris_version_2020} \\
                      \cline{2-4}

                      & Water restriction & precipitation \& potential evapotranspiration & \citet{harris_version_2020} \\
                      \cline{1-4}
\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & Sect. \ref{sec:irrigation} \\ 
                      \cline{2-4}

                      & Soil disturbance & tillage & Sect. \ref{sec:irrigation} \\
 \bottomhline
 \end{tabular}
 \belowtable{}
 \label{tab:datasourcedecay}
 \end{table*}

@@

with \(A_{lu}\) being the land-use type specific areas, \(AR_{lu}\) and \(AE_{lu}\) the area reduction resp. area expansion of the two land-use types. Data sources and methodology on land-use states and changes are described in Sect. \ref{sec:landuse}.

@@

>>>>

%% Journal abbreviations (please use the same for preprints and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% EGUsphere (egusphere) | This is only for EGUsphere preprints submitted without relation to an EGU journal.
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Polarforschung - Journal of the German Society for Polar Research (polf)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Safety of Nuclear Waste Disposal (sand)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

@@

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
% % % From pandoc table feature
% \usepackage{longtable,booktabs,array}
% % \usepackage{calc} % for calculating minipage widths
% % Correct order of tables after \paragraph or \subparagraph
% \usepackage{etoolbox}
% \makeatletter
% \patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
% \makeatother
% % Allow footnotes in longtable head/foot
% \IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
% \makesavenoteenv{longtable}
% 
% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
\begin{document}

@@

\begin{abstract}
Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth. The first meter of the Earth's soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere. SOC has been depleted by anthropogenic land-cover change and agricultural management. However, the latter has so far not been well represented in global carbon stock assessments. While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale. Here we develop a spatially explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We combine it with a reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils. We estimate that due to arable farming, soils have lost around 34.6 GtC relative to a counterfactual hypothetical natural state in 1975. Within the period 1975--2010 this SOC debt continued to expand by 5 GtC (\(0.14\unit{GtC yr^{-1}}\)) to around 39.6 GtC. However, accounting for historical management led to 2.1 GtC less (\(0.06\unit{GtC yr^{-1}}\)) emissions than under the assumption of constant management. We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique. The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling.
\end{abstract}

@@

Soil organic carbon (SOC), the amount of organic carbon stored in the Earth's soil, constitutes the largest terrestrial organic carbon pool. It exceeds the carbon in the atmospheric and vegetation pools multiple times \citep{batjes_total_1996}. Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere \citep{frielingstein_glob_carbon_2020, minasny_soil_2017}. The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile \citep{batjes_total_1996, sanderman_soil_2017}.

Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time. Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content \citep{batjes_harmonized_2016, hengl_soilgrids250m_2017, fao_global_2018}. However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis. Compared to climatic and geological driving forces, human interventions alter terrestrial carbon pools over much shorter time scales and are currently one of the most dominant drivers of SOC changes on managed land \citep{hansis_relevance_2015, bastos_comparison_2021}.

The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions, see Pongratz et al.~\citeyearpar{pongratz_terminolgy_2014}), which is the amount of organic carbon soils have lost under cultivation compared to a hypothetical potential natural vegetation state. Sanderman et al.~\citeyearpar{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter \citep{lal_world_2001}.

Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs), Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions \citep{frielingstein_glob_carbon_2020}. While providing estimates of the magnitude of SOC losses due to land-cover change, most models lack a detailed representation of agricultural management. Earlier DGVM- and ESM-based assessments only considered changes in land cover, but ignored the removal of biomass at harvest \citep{strassmann_simulating_2008, betts_climate_2015}. BKMs are designed to estimate LUC related emissions but often ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization and N deposition. Whereas BKM have largely improved in estimating additional emissions from wood harvest and shifting cultivation, state of the art models do not consider impacts of varying agricultural management \citep{frielingstein_glob_carbon_2020, houghton_carbon_2012, hansis_relevance_2015, bastos_comparison_2021}.

Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance \citep[e.g.~][]{bondeau_modelling_2007, lindeskog_implications_2013}. Pugh et al.~\citeyearpar{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historical management data. They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historical LUC and the slow decline of SOC under cropland before a new equilibrium is reached.

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons \citep{portmann_mirca2000global_2010}, multiple cropping systems \citep{waha_multiple_2020}, irrigation systems \citep{jagermeyr_water_2015}) or as stylized scenarios \citep[e.g.~][]{pugh_simulated_2015, lutz_simulating_2019}. Herzfeld et al.~\citeyearpar{herzfeld_soc_2021} account for historical changes in fertilizer and manure inputs, residue removal rates and tillage systems and report SOC losses from cropland expansion over the period from 1700--2018 of 215 GtC. Within their stylized future management scenarios they find that none of the management aspects considered (residue management, no-tillage) can create a net carbon sink on current cropland areas under future climate change.

More data sets on spatially explicit agricultural management time series with global coverage are becoming available (e.g.~on tillage systems, see \citep{porwollik_generating_2019, prestele_spatially_2018}) and modeling approaches are increasingly being developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019, minoli_modelling_2019}), but have --- to our knowledge --- not yet found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.

Field-scale models \citep{del_grosso_simulated_2001, coleman_simulating_1997, smith_estimating_2010, taghizadeh-toosi_c-tool_2014} are able to better account for historical agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge \citep{morais_detailed_2019}.

Managed soils have been increasingly studied not only for their carbon emitting behavior, but also because of their capacity to re-store carbon (soil carbon sequestration (SCS) techniques). However, assessing SCS dynamically considering the interdependency with environmental, social and economic sustainability targets has been difficult so far, as integrated assessment models (IAMs) \citep{popp_land-use_2016, rogelj_scenarios_2018} have not integrated soil management into their mitigation pathways. More detailed process-based models are typically computationally too demanding to be integrated into optimization-based IAMs. Better accounting for soil carbon management in IAMs thus requires a light-weight model suitable for iterative modeling with detailed options to represent agricultural soil management.

The objectives of our study are (1) to develop a reduced-complexity SOC model able to account for SCS in IAM frameworks; (2) to create a comprehensive data set of the global gridded management time series, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices; and (3) to provide global as well as spatially explicit SOC and SOC debt estimates that consider spatially explicit and time-variant agricultural management.
We decompose the contribution of different management activities through a scenario analysis, identifying the most impacting management decisions for SOC development. Moreover, we compare our model performance against other SOC stock and SOC emission estimates, to evaluate the suitability of this reduced-complexity approach for integration into IAM modeling.
\newpage

@@

In Sect. \ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail in the refinement of the IPCC guidelines vol.~4 Chapter 5 on ``Cropland'' \citep{ogle_cropland_in_ipcc_2019}. We additionally describe how we configured and extended the Tier 2 modeling approach \citep[for model code see][]{karstens_mrsoil_2020}. In Sect. \ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019}.
Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices \citep[for model code see][]{bodirsky_mrcommons_2020}. In Sect. \ref{sec:scenarios} we define the management scenarios used to analyze the role of different management aspects in historical cropland SOC dynamics.

\hypertarget{sec:carbonbudget}{%
\subsection{SOC stocks and stock changes following the Tier 2 modeling approach}\label{sec:carbonbudget}}

Following the Tier 2 modeling approach of the refinement of the IPCC guidelines vol.~4 Chapter 5 on ``Cropland'' \citep{ogle_cropland_in_ipcc_2019}; referred to as \textit{Tier 2 modeling approach} in the following), we estimate SOC stocks and their change over time for cropland at half-degree resolution from 1975 to 2010. We restrict our analysis to the first 0-30 cm of the soil profile. Moreover, we assume the current SOC state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
Therefore, we take the following three steps for each year of our simulation period:
(1) We calculate annual land-use type-specific steady states and decay rates for \(SOC\) stocks (Sect. \ref{sec:steadystates});
(2) we account for land conversion by transferring \(SOC\) from and to natural vegetation (Sect. \ref{sec:carbontransfer}),
(3) we estimate \(SOC\) stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \ref{sec:totalsoc}).
To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \ref{sec:initsoc}).

@@

In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{\mathrm{eq}}\) are given by
\begin{equation}
SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
\label{eq:inoutflow}
\end{equation}
with \(C^{\textrm{in}}\) being the carbon inputs to the soil, \(k\) denotes the soil organic carbon decay rate. This equation is valid for all grid cells \(i\) and all years \(t\). We use the Tier 2 modeling approach for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model \citep{parton_analysis_1987}. Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).

Carbon inputs for cropland are below- and above-ground crop residues left or returned to the field (see Sect. \ref{sec:residues}) and manure inputs (see Sect. \ref{sec:livstmanure}); for natural vegetation, litterfall including fine root turnover \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil. Following the IPCC guidelines \citep{ogle_cropland_in_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content. For each component the sum of all carbon input sources is allocated to the respective \(SOC\) sub-pools via transfer coefficients. This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools.

Whereas residue and manure default lignin and nitrogen fractions are given by the IPCC guidelines \citep{ogle_cropland_in_ipcc_2019}, we use plant-functional type and plant-organ specific parameterization for the natural litterfall. Global distribution of plant-functional types is given by \citep{schaphoff_lpjml4_2018} as well as separation of litter into leaf, fine root and wood litter compartments excluding litter biomass burnt in wild fires. Leaf litter parameters are given by Brovkin et al.~\citeyearpar{brovkin_plant-driven_2012}, fine root to leaf litter lignin ratio by Guo et al.~\citeyearpar{guo_leaf_root_decompostion_2021}, lignin content of wood litter by Rahman et al.~\citeyearpar{rahman_lignin_2013} and nitrogen content scaling factors for leaf to fine roots and leaf to wood litter by von Bloh et al.~\citeyearpar{vBloh_lpjml5_2018}.
Data sources for all considered carbon inputs as well as for lignin and nitrogen parameterization are listed in Table \ref{tab:datasourceinputs}.

  \begin{table*}[h]
  \caption{type and data sources for carbon inputs and parameterization to different land-use types }
  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
  \tophline
  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
  \middlehline
  \multirow{3}{*}{cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                      \citet{faostat_faostat_2016}, \\
                                                                      \citet{schaphoff_lpjml4_2018},\\
                                                                      \citet{weindl_livestock_2017} \end{minipage}} & 
                         \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                         LG:C generic values according to Table 5.5B,\\
                         5.5C from IPCC \citep{ogle_cropland_in_ipcc_2019},\\
                         crop-specific N:C from \citet{bodirsky_n2o_2012}
                                                                      \end{minipage}} \\
                            & below-ground residues, &  &  \\
                            & manure                &  &  \\ 
                            \hline
  natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
                      & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
                          leaf N and LG concentration from \citet{brovkin_plant-driven_2012}, 
                          root to leaf litter LG ratio \citet{guo_leaf_root_decompostion_2021},  
                          lignin content of wood litter \citet{rahman_lignin_2013} and 
                          nitrogen scaling factors for leaf to root and wood litter 
                          from \citet{vBloh_lpjml5_2018} \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \label{tab:datasourceinputs}
 \belowtable{}
 \end{table*}

The sub-pool specific decay rates \(k_{sub}\) are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space \(i\) and time \(t\). Following the Tier 2 modeling approach \citep{ogle_cropland_in_ipcc_2019}, we consider temperature \(temp\), water \(wat\), sand-fraction \(sf\), and tillage \(till\) effects to account for spatial and temporal variation of decay rates. Thus, \(k_{sub}\) rates are given by

@@

For natural vegetation, we assume rainfed and non-tilled conditions, whereas for cropland, we distinguish the effect of different tillage (see Sect. \ref{sec:tillage}) and irrigation (see Sect. \ref{sec:irrigation}) practices on decay rates. We calculate area-weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation practices. Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equations 5.0B--5.0F in Ogle et al.~\citeyearpar{ogle_cropland_in_ipcc_2019}.

 \begin{table*}[h]
 \caption{type and data sources for carbon inputs to different land-use types}
 \begin{tabular}{l l l l}
 \tophline
  land-use types   & type of decay driver & parameter use to represent driver & data source \\
 \middlehline
 \multirow{2}{*}{all} & soil quality & sand fraction of the first 0-30 cm 
                                     & \citet{hengl_soilgrids250m_2017} \\
                      \cline{2-4}

                      & mircobial activity & air temperature & \citet{harris_version_2020} \\
                      \cline{2-4}

                      & soil moisture & precipitation \& potential evapotranspiration & \citet{harris_version_2020} \\
                      \cline{1-4}
\multirow{2}{*}{\begin{minipage}[t]{0.15\columnwidth}\raggedright\strut cropland\\(additionally)\strut\end{minipage}} & soil moisture*  & irrigation  & Sect. \ref{sec:irrigation} \\ 
                      \cline{2-4}

                      & soil disturbance & tillage & Sect. \ref{sec:tillage} \\
 \bottomhline
 \end{tabular}
 \belowtable{}
 \label{tab:datasourcedecay}
 \end{table*}

@@

with \(A_{lu}\) being the land-use type specific areas, \(AR_{lu}\) the area reduction and \(AE_{lu}\) the area expansion of the two land-use types. Data sources and methodology on land-use states and changes are described in Sect. \ref{sec:landuse}.

@@

\begin{table}[htbp]
  \centering
  \caption{Mapping of FAO crop production items \citep{faostat_faostat_2016} to LUH2 categories \citep{hurtt_harmonization_2020} and to crop groups used in this study.}
